A systematic comparison of interaction models for systems of cities

### 2018-09-23 Docs/CCS18/Communication/ccs2018.tex
v1 slides

\date{CCS 2018\\\smallskip
Thessaloniki\\\smallskip
September 25th 2018
}

@@

\sframe{Urban systems}{

@@

	%\includegraphics[width=\textwidth]{figures/}

	\footnotesize
\textit{Source: }

@@

% meso / micro

% macro

@@

\sframe{An evolutionary urban theory}{
\centering

@@

@@

\textbf{Research objective : }

@@

Comparison of three approaches based on the evolutionary urban theory \cite{pumain1997pour} capturing different dimensions of urban systems:
\begin{itemize}
	\item The Favaro-Pumain model for the diffusion of innovation \cite{favaro2011gibrat}
	\item The Marius model family based on economic exchanges \cite{cottineau2014evolution}
	\item An interaction model including physical transportation networks \cite{raimbault2018indirect}
\end{itemize}

@@

\sframe{Network interaction model}{

@@

\sframe{The Favaro-Pumain model}{

@@

@@

Harmonized dataset, explored in particular by \cite{pumain2015multilevel} (built by

@@

\sframe{Data preprocessing}{
\centering

Remove small cities (\cite{adam2006medium} for definition of medium-sized)

@@

\sframe{Stylized facts}{
\centering

@@

Computationally intensive: high-dimensional parameter space and possible spatial setup.

$\rightarrow$ use of grid computing, made smooth with the OpenMOLE software \url{https://next.openmole.org/}

@@

@@

@@

\sframe{Results: comparison of models on the different systems}{

@@

\sframe{Results: particular cases}{

@@

@@

$\rightarrow$

@@

$\rightarrow$

@@

% future perspective : towards integrative theories / coupling of these models for multi-dimensional grasp of urban systems (more work needed to understand model coupling and role of parameters in fit improvement)
% Compare these models on synthetic city systems / with suited indicators (here relatively data-driven approach)

$\rightarrow$

@@

$\rightarrow$

@@

$\rightarrow$ More elaborated method to compare models in a ``fair'' way (correcting for additional parameters, open question for models of simulation).

@@

$\rightarrow$ % model coupling / benchmarking

@@

\textbf{Open repository} (code, data and results) at\\\texttt{https://github.com/JusteRaimbault/UrbanGrowth}\\\medskip
\textbf{Acknowledgments} : I thank the \textit{European Grid Infrastructure} and its \textit{National Grid Initiatives} (\textit{France-Grilles} in particular) to give the technical support and the infrastructure.

@@

>>>>

\date{CCS 2018\\\smallskip
Thessaloniki\\\smallskip
September 24th 2018
}

@@

\sframe{The long time memory of urban systems}{

@@

	\includegraphics[width=0.52\textwidth]{figures/rotundagalerius.jpg}
	\hspace{0.2cm}
	\includegraphics[width=0.42\textwidth]{figures/whitetower.png}

	\footnotesize
\textit{Source: Wikipedia}

@@

% meso / micro - macro

\begin{columns}
	\begin{column}{0.5\textwidth}
		Macroscopic urban system
		\medskip
		%\centering
		\includegraphics[width=\textwidth]{figures/macrogrowth.png}
		\footnotesize
		\justify
		Raimbault, J. (2018). Indirect evidence of network effects in a system of cities. \textit{Environment and Planning B: Urban Analytics and City Science}, 2399808318774335.
		\nocite{raimbault2018indirect}
	\end{column}
	\begin{column}{0.5\textwidth}
		Mesoscopic territorial systems
		\medskip
		%\centering
		\includegraphics[width=\textwidth]{figures/mesogrowth.png}

		\footnotesize
		\justify
		Raimbault, J. (2018). Calibration of a density-based model of urban morphogenesis. PloS one, 13(9), e0203516.
		\nocite{raimbault2018calibration}
	\end{column}
\end{columns}

@@

\sframe{An evolutionary urban theory}{

From \cite{pumain1997pour} to \cite{pumain2018evolutionary}: systems of cities as co-evolutive systems in which interactions are crucial

\includegraphics[height=0.7\textheight]{figures/evolurbantheory.pdf}

@@

$\rightarrow$ several models in this context have been introduced, but never compared (e.g. in terms of explanative power)

$\rightarrow$ multidimensionality of urban systems and the potential complementarity between very different processes

@@

\bigskip
\bigskip

\textbf{Research objective : } \textit{Benchmark interaction models for systems of cities developed in the frame of the evolutionary urban theory, on several comparable systems of cities.}

@@

%\centering
\justify

Comparison of three approaches based on the evolutionary urban theory \cite{pumain1997pour} capturing different dimensions of urban systems:

\begin{itemize}
	\item The Favaro-Pumain model for the diffusion of innovation \cite{favaro2011gibrat}
	\item The Marius model family based on economic exchanges \cite{cottineau2014evolution}
	\item An interaction model including physical transportation networks \cite{raimbault2018indirect}
\end{itemize}

@@

\sframe{Description of models}{

@@

\begin{columns}
	\begin{column}{0.33\textwidth}
		\textbf{Network interaction model}

		\begin{itemize}
			\item Endogenous growth
			\item Interactions inducing growth through gravity potential
			\item Static physical network taken into account (geographical shortest path with topography)
		\end{itemize}

	\end{column}
	\vrule{}
	\begin{column}{0.33\textwidth}
		%\vspace{-1cm}
		\textbf{Favaro-Pumain model}

		\begin{itemize}
			\item Endogenous growth
			\item Innovation emerge and diffuse in cities
			\item Growth rates adapted according to utility of innovation and level of adaptation
		\end{itemize}

	\end{column}
	\vrule{}
	\begin{column}{0.33\textwidth}
		\textbf{Marius model}

		\begin{itemize}
			\item Cities produce economic goods
			\item Economic exchanges are estimated according to gravity flows
			\item Populations grow depending on final economic balances
		\end{itemize}

	\end{column}
\end{columns}

@@

%\centering

@@

\textbf{Dataset}

Harmonized dataset, explored in particular by \cite{pumain2015multilevel} (built in the context of Geodivercity ERC)

\textbf{Data preprocessing}

\begin{itemize}
	\item Remove small cities and medium-sized cities (\cite{adam2006medium} for definition of medium-sized) for scalability of simulations
	\item Shortest paths matrices (direct and taking into account topography, using the EEA world DEM) computed a priori and cached (\textit{no evolution of the network})
\end{itemize}

@@

\sframe{Stylized facts}{
%\centering

\textit{Population distributions mainly log-normal (75\% with KS-test, all with AIC); consistent hierarchies in time.}

\includegraphics[width=0.49\textwidth]{figures/popdistribs.png}
\includegraphics[width=0.49\textwidth]{figures/popranksize.png}

@@

\sframe{Stylized facts}{

\textit{Growth rates distributions evolve considerably in time}

\includegraphics[width=\textwidth]{figures/growthrates_byyear.png}

\sframe{Stylized facts}{

\textit{Correlation as a function of distance: long-range correlations in several systems}

\includegraphics[width=0.49\textwidth]{figures/corrsdist_US_quantiles21_periods5.png}
\includegraphics[width=0.49\textwidth]{figures/corrsdist_EU_quantiles21_periods5.png}

%\smallskip

%$\rightarrow$ \textit{non-linear models good candidates}

@@

Bi-objective calibration of 7 models for 7 city systems $\rightarrow$ use of genetic algorithms on grid, made smooth with the OpenMOLE software \url{https://next.openmole.org/}

@@

\textbf{Come to the Satellite on Wednesday, and apply to the summer school !} 
\\(\texttt{https://exmodelo.org/})

@@

\sframe{Comparison of models on all systems}{

\includegraphics[width=\textwidth]{figures/allmodels_allsystems.png}

\footnotesize
\textit{Systems driven by economic exchanges: EU, US, RU, ZA}

@@

\sframe{Indian urban system: direct interactions}{

\centering
\includegraphics[width=0.9\textwidth]{figures/IN_zoomed.png}

@@

\sframe{Brazilian urban system: multiple factors}{

\centering
\includegraphics[width=0.9\textwidth,height=0.8\textheight]{figures/BR_zoomed.png}

\footnotesize
\textit{Importance of topography; innovation processes mostly.}

\sframe{China: a tight competition}{

\centering
\includegraphics[width=0.9\textwidth,height=0.8\textheight]{figures/CN_zoomed.png}

\footnotesize
\textit{No clear best model: other processes in play ? (strong top-down planning)}

@@

\textit{Comparing Gibrat and Network model by correcting for the number of parameters \cite{raimbault2018indirect}: additional parameters actually improve the fit (but not always)}

% gib - intgib
\begin{tabular}{|c|c|c|c|c|}
\hline
System & $\Delta AIC$ (log) & $\Delta BIC$ (log) & $\Delta AIC$ (mse) & $\Delta BIC$ (mse) \\\hline
ZA & 33.82658 & 29.40496 &  -838.07184 &  -842.49347\\
CN & 20.67713 & 16.78273 & 13.40767 & 9.51327\\
BR & -263.27003 & -267.71575 & -407.55220 & -411.99792\\
IN & -85.01949 & -89.24206 & -188.16703 & -192.38960\\
RU & 50.60535 & 46.24233 & 104.94806 &  100.58503\\\hline
\end{tabular}

% marius - mariusrestr

@@

$\rightarrow$ Complementarity of economic, innovation and direct interaction processes

@@

$\rightarrow$ High dimensionality of urban systems

@@

% future perspective : towards integrative theories / coupling of these models for multi-dimensional grasp of urban systems (more work needed to understand model coupling and role of parameters in fit improvement)
% Compare these models on synthetic city systems / with suited indicators (here relatively data-driven approach)
% - nonstat in time for some city systems ?
% - calib at the second order on correlation curves (-> new methodo approach - link with correlated synthetic data !) (many things are connected)
% saturation of models ?
% Rq : issue in calib algos (restricted vs full models : strange behavior ex the bifurcation in intgib % gibrat

$\rightarrow$ Still points on robustness to be investigated: influence of stochasticity, convergence of GA, saturation of parameters, number of cities.

@@

$\rightarrow$ Towards integrated models coupling these different components ?

@@

$\rightarrow$ Compare on synthetic systems of cities with appropriate indicators \cite{raimbault2018unveiling}.

$\rightarrow$ Calibrate at the second order (correlations), non-stationary in time.

$\rightarrow$ More elaborated method to compare models in a ``fair'' way (correcting for additional parameters, open question for models of simulation): model reduction ?

@@

$\rightarrow$ First step towards systematic benchmarks and multi-modeling. \textbf{Need for more systematic model exploration.} % model coupling / benchmarking

$\rightarrow$ Model integration and multi-scalarity ? \textbf{Need for more integrated models.}

@@

\textbf{Open repository} at \texttt{https://github.com/JusteRaimbault/UrbanGrowth}\\\smallskip
\textbf{Acknowledgments}: thanks to the \textit{EGI} for access to the infrastructure.

@@

Genetic algorithm NSGA2:

\begin{itemize}
	\item Around 40000 generations (4e6 model runs)
	\item Population $\mu = 100$
	\item Convergence tested with hypervolume relative variation
\end{itemize}
